---
title: "RepData_PeerAssessment1"
author: "Olusegun"
date: "4/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Loading and preprocessing the data

**1.Load the data (i.e. read.csv())**
```{r}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"./Activity_monitoring_data.zip")
unzip("./Activity_monitoring_data.zip", exdir = "./data_AMD")
activity <- read.csv("./data_AMD/activity.csv")
```

**2.Process/transform the data (if necessary) into a format suitable for your analysis**
```{r}
activity$date <- as.POSIXct(activity$date)
```

## What is mean total number of steps taken per day?

**1.Make a histogram of the total number of steps taken each day**
```{r}
sum_steps <- aggregate(steps~date,activity,sum)
hist(sum_steps$steps, main = "Histogram of the total number of steps taken each day", 
     xlab = "Total number of steps per day")
```

**2.Calculate and report the mean and median total number of steps taken per day**
```{r}
mean_sum_steps <- mean(sum_steps$steps)
median_sum_steps <- median(sum_steps$steps)
```
mean = `r mean_sum_steps`

median = `r median_sum_steps`

## What is the average daily activity pattern?

**1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**
```{r}
time_series <- aggregate(steps~interval,activity,mean)
plot(time_series, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Average number of steps taken")
```

**2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
```{r}
time_series[which.max(time_series$steps),][[1]]
```

## Imputing missing values

**1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**
```{r}
sum(is.na(activity))
```
There are `r sum(is.na(activity))` missing values in the dataset.

**2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**

Here I use the mean of 5-minute interval to fill in the values of the missing values.

**3.Create a new dataset that is equal to the original dataset but with the missing data filled in.**
```{r}
activity_narm <- activity
activity_narm <- merge(activity_narm,time_series,by="interval")
activity_narm <- activity_narm[order(activity_narm$date),]
activity_narm$steps.x[which(is.na(activity_narm$steps.x))]<-activity_narm$steps.y
activity_narm$steps.y <- NULL
colnames(activity_narm)[2] <- "steps"
sum(is.na(activity_narm))
```

**4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.**
```{r}
sum_activity_narm <- aggregate(steps~date,activity_narm,sum)
hist(sum_activity_narm$steps, main = "Histogram of the total number of steps taken each day", 
     xlab = "Total number of steps per day")
mean_sum_activity_narm <- mean(sum_activity_narm$steps)
median_sum_activity_narm <- median(sum_activity_narm$steps)
```
**Do these values differ from the estimates from the first part of the assignment?**
After replacing the mean is the same but the median is a little bit different
difference in mean = `r abs(mean_sum_activity_narm-mean_sum_steps)`
difference in median = `r abs(median_sum_activity_narm-median_sum_steps)`

**What is the impact of imputing missing data on the estimates of the total daily number of steps?**
The impact will be minute, since we have no difference in mean and slight difference in median

## Are there differences in activity patterns between weekdays and weekends?

**1.Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.**
```{r}
week_day <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity_narm$days <- factor(weekdays(activity_narm$date) %in% week_day, levels = c(TRUE,FALSE), labels = c("weekday","weekend"))
```

**2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). **
```{r}
time_series_2 <- aggregate(steps~interval+days,activity_narm,mean)
library(lattice)
xyplot(steps ~ interval | days, time_series_2, type = "l", layout = c(1, 2), 
     xlab = "Interval", ylab = "Number of steps")
```